---
title: "Lab 6 Penguins"
author: "Tyanna Bui"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Loading packages and read in the data
```{r, include = F}
library(tidyverse)
library(car)
library(onewaytests)

pen = read_csv("penguins.csv")
```

Our question today is... is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipeligo?

Maybe I can use a one-way ANOVA... assuming that
- Observations are independent
- Equal variance (Levene's test)
- Normality (histograms/qqplots)

```{r}
counts = with(pen, table(species, sex)) # using with(dataset, table with wanted colms)

pen2 = pen %>% 
  mutate(common_name = case_when(species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap", species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie", species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo")) %>% 
  select(-species) %>% # removing species column
  filter(sex == "FEMALE")

pen2
```

Exploratory graphs and Levene's test for equal variances

```{r}
# creating histogram
his = ggplot(pen2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) 

his # do these look like their means differ based on this? YES

# creating qqplot
qq = ggplot(pen2, aes(sample = flipper_length)) +
  geom_qq(aes(color = common_name))

qq # based on qq & hist, obs for ea group appear normally distributed. n > 30 so makes sense it's normally distributed by CLT.
```

Evaluate variances and perform a Levene's Test for equal variances
```{r}
# If largest sample variance is < 4x greater tha nthe smallest sample variance, the nthese are "close enough" to consider variances equal

variance = pen2 %>% 
  group_by(common_name) %>% 
  summarize(variance = var(flipper_length))
variance # unit is obs. unit^2... largest variance 33.1 is /> 4x greater than smallest variance (15.2)

# Levene's Test
# leveneTest function in 'car' package
levenepen = leveneTest(pen2$flipper_length ~ pen2$common_name)
levenepen

# Ho: Variances are =
# Ha: Variances are /= equal

# p value = 0.05782 --> no sig diff in variances across species
```

One-way ANOVA for penquin flipper lengths by species (common_name)
```{r}
# anova
anovatest = aov(pen2$flipper_length ~ pen2$common_name)
summary(anovatest) # to get full summary of anova outcome

# Ho: Means across ALL GROUPS are =
# Ha: At least 2 samples differ
# p value is almost 0... therefore must reject null and find out which sample is different from the rest, or if all of them are different?
# ----------------------------
# post-hoc test by Turkey's HSD
penph = TukeyHSD(anovatest)
penph

# Ho: No sig diff in pairwise means
# Ha: Sig diff in pairwise means
# p values indicate that all pairwise means are sig diff from each other
```
Mean flipper lengths of femaile penguins differed significantly for all species studied (one-way anova, F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05).

Note: aov.test funcrion in the `onewaytests` package DOES allow for inline references.
